---
title: "Darwin Core Mapping"
subtitle: "For: MijnVISmaat - Exotic fish occurrences in Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(lubridate)      # date transformation
```

# Read source data

```{r}
input_data <- read_delim(
  file = here::here("data", "raw", "MVM_hengelvangsten_Vlaanderen.csv"),
  ";",
  escape_double = FALSE,
  trim_ws = TRUE,
  guess_max = 5000
)
```

Preview data:

```{r}
input_data %>% head()
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```


# Darwin Core mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
occurrence <- input_data
```

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "Royal Dutch Angling Association")
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use")
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/6va6gm")
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "Royal Dutch Angling Association")
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "MijnVISmaat - Exotic fish occurrences in Belgium")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = ID) 
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = "MijnVISmaat")
```

### individualCount

```{r}
occurrence %<>% mutate(dwc_individualCount = "1") 
```

### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_individualCount = "present")
```

## samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = "fishing rod catch") 
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = format_ISO8601(
  # Parse as # 21-09-2014 16:30:00
  parse_date_time(paste(datum_dd_mm_jjjj, tijdconversie_uu_mm), "%d%m%y %h%m%s"),
  # Convert to iso 2014-09-21T16:30
  precision = "ymdhm"
))
```

### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe") 
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = "BE") 
```

### verbatimLocality

```{r}
occurrence %<>% mutate(dwc_verbatimLocality = Visstek)
```

### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.5f",
  round(lat_epsg_4326, digits = 5)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.5f",
  round(lon_epsg_4326, digits = 5)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84")
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = "30")
```

### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = acceptedName)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = case_when(
  scientificName == "Cyprinus carpio x Carassius auratus" ~ "hybrid",
  TRUE ~ "species"
))
```

### vernacularName

```{r}
occurrence %<>% mutate(dwc_vernacularName = Vissoort)
```

# nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

## Remove rows with non invasive fish

```{r}
occurrence %<>%
  filter(scientificName != 'Leuciscus idus (Linnaeus, 1758)') %<>%
  filter(scientificName != 'Silurus glanis Linnaeus, 1758')
```

## Post-processing

```{r}
occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("data", "processed", "occurrence.csv"), na = "")
```

# Data quality tests

```{r}
parsed_names <- input_interim %>%
  distinct(acceptedName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function

write_csv(parsed_names, here::here("data", "interim", "species.csv"))
```
