---
title: "Darwin Core Mapping"
subtitle: "For: MijnVISmaat - Exotic fish occurrences in Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # date transformation
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv



```{r}

input_interim <- read_delim(file = here::here("data", "raw", "MVM_hengelvangsten_Vlaanderen.csv"), ";" , escape_double = FALSE , trim_ws = TRUE, guess_max = 5000  )
##input_interim <- read_csv2(file = here::here("data", "RAW", "MVM_hengelvangsten_Vlaanderen.csv"))                        

```

Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```


```{r}
input_interim %>% head(n = 5)
```

# georefernce source data

## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates

We want to keep the original verbatim Coordinates
The original X,Y will be used for creating decimalLatitude and decimalLongitude

```{r}
input_interim %<>% mutate(decimalLongitude = `Lon (EPSG: 4326)`) %>%
                   mutate(decimalLatitude = `Lat (EPSG: 4326)`)
```
### Create a spacial dataframe

Create spatial dataframe and define the CRS (31370 is Belgian Lambert 72)(crs = 28992 is the Dutch system used here)
We indicate the x,y columns as coordinates and define the coordinate system


```{r}
#input_interim = st_as_sf(input_interim, coords = c("Longitude","Latitude"), crs = 28992 , agr = "constant" )
#input_interim %>% head(n = 5)
```


## TRansform the data in WGS84

Now we transform the Lambert 72 in WGS84

```{r}
#input_interim <- st_transform(input_interim, crs = 4326)
#input_interim %>% head(n = 5)

```

### Create a dataframe with column decimalLatitude & decimalLongitude

In order to add the decimal coordinates into the original df we create an x_y dataframe

```{r}
#x_y <- st_coordinates(input_interim)
#x_y %>% head(n = 5)

#write.csv(x_y, "../data/interim/x_y.csv", append = FALSE, na = "",  row.names = FALSE, fileEncoding = "UTF-8")
```
# create extra dataframe to work with 

```{r}
#input_interim_df <- input_interim
```


# remove geometry from dataframe 'input_interim_df'

we remove all geometry from the spacial dataframe, to create a normal df

```{r}
##input_interim_df <- st_geometry(NULL)

#st_geometry(input_interim_df) <- NULL

#input_interim_df %>% head(n = 5)

```


### bind columns X_y and input_interim_df

join both df's to have the decimalLat & Lon

```{r}

#input_interim_df = bind_cols(input_interim_df,x_y, .id = NULL)

#input_interim_df %>% head(n = 5)

```

### Integrate x-y coordinates in original dataframe

We can safely substitute input_interim 


```{r}
#input_interim <- input_interim_df
```



# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}

head(occurrence, n =5)
##occurrence %<>% select(-c(kind_nl,day, month, year))
```



# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "my_doi") # e.g. "to compplete"
```

### type

```{r}
occurrence %<>% mutate(type = "Event")
```


### language

```{r}
occurrence %<>% mutate(language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(rightsHolder = "Royal Dutch Angling Association") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(institutionCode = "Royal Dutch Angling Association") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(datasetName = "MijnVISmaat app, alien fish occurrences in Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basis of Record

```{r}
occurrence %<>% mutate(basisOfRecord = "HumanObservation")
```
### informationWithHeld

### dataGeneralizations

### occurrenceID

**This need to be fixed**

```{r}
occurrence %<>% mutate(occurrenceID = ID) 
#                mutate(occurrenceID = str_c("UVW:", occurrenceID))
  
```


### recordedBy

```{r}
occurrence %<>% mutate(recordedBy = 'MijnVismaat')
```


### individualCount

```{r}
occurrence %<>% mutate(individualCount = '1') 
```

### oragnismQuantity

### organismQuentityType

### sex

```{r}
#occurrence %<>% mutate(sex = recode(Gender,
#"Muskusrat" = "Muskusrat",
#                   "jong"  = "juvenile",
#                   "oud" = "adult",
#                   "Onbekend" = "",
#                                   ))
```


### lifeStage

```{r}
#occurrence %<>% mutate(lifeStage = recode(Age,
#"Muskusrat" = "Muskusrat",
#                   "moer"  = "female",
#                   "ram" = "male",
#                   "Onbekend" = "",
#                                   ))
```



### behavior


### occurrenceRemarks

```{r}
# occurrence %<>% rename(occurrenceRemarks = 'action_en')
```
## samplingProtocol

```{r}
occurrence %<>% mutate(samplingProtocol = 'fishing rod catch') 
 

 # extract follow ups to new column
# mutate(follow_up = str_extract(samplingProtocol, "Follow-up=[0-9]*")) %>%
 # remove follow ups from samplingProtocol
# mutate(samplingProtocol = str_remove(samplingProtocol, "Follow-up=[0-9]*")) %>%
 # remove leading ", " if present
# mutate(samplingProtocol = str_remove(samplingProtocol, "^,\\s")) %>%
 # remove trailing ", " if present
# mutate(samplingProtocol = str_remove(samplingProtocol, ",\\s$"))%>%
 # rename follow up
# rename(samplingEffort = "follow_up") 
 # replace follow-up with times visited
 # mutate(samplingEffort = str_replace(samplingEffort, "Follow-up=*", "times-visited="))
```
### samplingEffort

```{r}
# occurrence %<>%
# extract follow ups to new column
# mutate(samplingEffort = str_replace(samplingEffort, "Follow-up=*", "times-visited=")) 


```


### eventDate


```{r}
# occurrence %<>% rename(eventDate = 'CreatedOn')



```
```{r}
occurrence %<>% mutate(eventDate = dmy(`Datum (dd-mm-jjjj)`)) 
```
```{r}
occurrence %<>% mutate(year = year(eventDate)) %<>%
                mutate(month = month(eventDate)) %<>%
                mutate(day = day(eventDate)) %<>%
                mutate(eventDate = date(eventDate))
                
```



# Location



```{r}

occurrence %<>% mutate(geodeticDatum = "WGS84") %>%
                mutate(coordinateUncertaintyInMeters = "30") %>%
                # mutate(verbatimCoordinateSystem = "Rijksdriehoekscoördinaten") %>%
                mutate(countryCode = "BE")  %>%            
                mutate(continent = "Europe")


```


```{r}

head(occurrence, n = 5)
occurrence %<>%
  # mutate(verbatimLongitude = round(verbatimLongitude)) %>%
  # mutate(verbatimLatitude = round(verbatimLatitude)) %>%
   mutate(decimalLongitude = round(decimalLongitude, digits = 5)) %>%
   mutate(decimalLatitude = round(decimalLatitude, digits = 5)) 
   
```


```{r}
occurrence %<>%   
   mutate(decimalLatitude = as.character(format(decimalLatitude, nsmall = 5))) %>%
   mutate(decimalLongitude = as.character(format(decimalLongitude, nsmall = 5)))

```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "NL") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
#occurrence %<>%
#  mutate(municipality = str_remove(municipality,"\\(BE\\)"))
```

###LocationID

```{r}
#occurrence %<>% mutate(locationID = SubAreaName)
```

### locality

```{r}
occurrence %<>% mutate(verbatimLocality = Visstek)
```





### verbatimcoordinates

```{r}
#occurrence %<>% mutate(verbatimCoordinates = HourSquareName)
```



### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem


```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Amersfoort/RD new") # 
```

### verbatimSRS

```{r}
occurrence %<>% mutate(verbatimSRS = "EPSG:4326")
```

## decimalLatitude
## decimalLongitude
## geodeticDatum
## coordinateUncertaintyInMeters
## georeferenceRemarks

# Occurrence

###identifiedBy
```{r}
occurrence %<>% mutate(identifiedBy = 'MijnVISmaat, Royal Dutch Angling Association')
```

### scientificName 

```{r}
head(occurrence, n=5)
occurrence %<>% mutate(scientificName = acceptedName)

# recode(CatchTypeName,
#                   "Muskusrat" = "Ondatra zibethicus",
#                   "Beverrat"  = "Myocastor coypus",
#                   "Muskusrat ram oud (>1jr)" = "Ondatra zibethicus",
#                   "Muskusrat ram jong (<1jr)" = "Ondatra zibethicus",
#                   "Muskusrat moer oud (>1jr)" = "Ondatra zibethicus",
#                   "Muskusrat moer jong (<1jr)" = "Ondatra zibethicus",
#                                       ))
            


```
### kingdom
```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the dataset:

```{r}
parsed_names <- occurrence %>%
  distinct(scientificName) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Preview data:

```{r}
parsed_names %>% head(n = 62)
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

#verbatimScientificName


```{r}
occurrence %<>% mutate(vernacularName = Vissoort)
```
#taxonRank
```{r}
occurrence %<>% mutate(taxonRank = case_when(scientificName == "Cyprinus carpio x Carassius auratus" ~ "hybrid",
                                             TRUE ~ "species"))


#taxon %<>% mutate(nomenclaturalCode =case_when(kingdom == "Animalia" ~ "ICZN",kingdom == "Plantae" ~ "ICBN"  )) 



```


# nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```


### OccurrenceStatus present/absent

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(occurrenceStatus = case_when(individualCount > 0 ~ "present",
                              individualCount == 0 ~ "absent"
                                 )
                                )

```

## remove rows with non invasive fish

```{r}

occurrence %<>%
        filter(scientificName != 'Leuciscus idus (Linnaeus, 1758)') %<>%
        filter(scientificName != 'Silurus glanis Linnaeus, 1758')

        
  
  
  
 


```



## Post-processing


```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
occurrence %<>% select(-c(year, month, day, Foto,`Datum (dd-mm-jjjj)`, `Tijdconversie (uu:mm)`,Visstek, Vissoort, `Lat (EPSG: 4326)`, `Lon (EPSG: 4326)`, acceptedName, Bron ))  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```


define the order of the output columns

```{r}
col_order <- c( "ID","type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","samplingProtocol","eventDate","continent","countryCode","verbatimLocality"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","taxonRank","nomenclaturalCode"
               ) 
               # still needed for this dataset ,"occurrenceID"
               # removed for this dataset,"occurrenceRemarks","municipality","samplingEffort"
               # removed for this dataset, "verbatimLatitude","verbatimLongitude","verbatimCoordinates"
               # ,"verbatimCoordinateSystem","verbatimSRS" ,"scientificNameAuthorship"
occurrence <- occurrence[, col_order]
```


Preview data:

```{r}
occurrence %>% head()

```

Save to CSV:

```{r}
write.csv(occurrence, "../data/processed/occurrence.csv", append = FALSE, na = "",  row.names = FALSE, fileEncoding = "UTF-8")
```


```{r}
#write_csv(occurrence, here("data", "processed", "occurrence.csv"), na = "")
#write_csv(occurrence, here("data", "processed", "occurrence.csv"), append = FALSE, na = "")

```


# data quality tests

```{r}

parsed_names <- input_interim %>%
  distinct(acceptedName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function

head(parsed_names, n = 30)

write.csv(parsed_names, "../data/interim/species.csv", append = FALSE, na = "",  row.names = FALSE, fileEncoding = "UTF-8")

```

