---
title: "Darwin Core Mapping"
subtitle: "For: MijnVISmaat - Exotic fish occurrences in Belgium"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(lubridate)      # date transformation
```

# Read source data

```{r}
input_data <- read_delim(
  file = here::here("data", "raw", "MVM_hengelvangsten_Vlaanderen.csv"),
  ";",
  escape_double = FALSE,
  trim_ws = TRUE,
  guess_max = 5000
)
```

Preview data:

```{r}
input_data %>% head()
```

# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```


# Darwin Core mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
occurrence %<>% mutate(datasetID = "my_doi") # e.g. "to compplete"
```

### type

```{r}
occurrence %<>% mutate(type = "Event")
```

### language

```{r}
occurrence %<>% mutate(language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(rightsHolder = "Royal Dutch Angling Association") # e.g. "INBO"
```

### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(institutionCode = "Royal Dutch Angling Association") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(datasetName = "MijnVISmaat app, alien fish occurrences in Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

### basis of Record

```{r}
occurrence %<>% mutate(basisOfRecord = "HumanObservation")
```

### occurrenceID

```{r}
occurrence %<>% mutate(occurrenceID = ID) 
#                mutate(occurrenceID = str_c("UVW:", occurrenceID))
  
```

### recordedBy

```{r}
occurrence %<>% mutate(recordedBy = 'MijnVismaat')
```

### individualCount

```{r}
occurrence %<>% mutate(individualCount = '1') 
```


```{r}
```

## samplingProtocol

```{r}
```

### eventDate

```{r}
```

### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
```

### locality

```{r}
occurrence %<>% mutate(verbatimLocality = Visstek)
```


```{r}
```


```{r}
```


```{r}
occurrence %<>% mutate(verbatimSRS = "EPSG:4326")
```

## decimalLatitude
## decimalLongitude
## geodeticDatum
## coordinateUncertaintyInMeters
## georeferenceRemarks

# Occurrence

###identifiedBy
```{r}
occurrence %<>% mutate(identifiedBy = 'MijnVISmaat, Royal Dutch Angling Association')
```

### scientificName 

```{r}
occurrence %<>% mutate(kingdom = "Animalia")
```


```{r}
```


```{r}
```


```{r}
occurrence %<>% mutate(vernacularName = Vissoort)
```

# nomenclaturalCode

```{r}
```

## remove rows with non invasive fish

```{r}
occurrence %<>%
```

## Post-processing

```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
occurrence %<>% select(-c(year, month, day, Foto,`Datum (dd-mm-jjjj)`, `Tijdconversie (uu:mm)`,Visstek, Vissoort, `Lat (EPSG: 4326)`, `Lon (EPSG: 4326)`, acceptedName, Bron ))  #remove collection columns
          ##  rename(scientificName = verbatimScientificName) 
```


define the order of the output columns

```{r}
col_order <- c( "ID","type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","occurrenceStatus","samplingProtocol","eventDate","continent","countryCode","verbatimLocality"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","taxonRank","nomenclaturalCode"
               ) 
               # still needed for this dataset ,"occurrenceID"
               # removed for this dataset,"occurrenceRemarks","municipality","samplingEffort"
               # removed for this dataset, "verbatimLatitude","verbatimLongitude","verbatimCoordinates"
               # ,"verbatimCoordinateSystem","verbatimSRS" ,"scientificNameAuthorship"
occurrence <- occurrence[, col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("data", "processed", "occurrence.csv"), na = "")
```

# Data quality tests

```{r}
parsed_names <- input_interim %>%
  distinct(acceptedName)  %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function

write_csv(parsed_names, here::here("data", "interim", "species.csv"))
```
